Covid-19 infection and attributable mortality in residents of UK care homes: Cohort study using electronic records (March-June 2020)

### 2020-07-08 manuscript/manuscript.Rmd
excess mortality

---
title: "Covid-19 infection and attributable mortality in residents of UK care homes: Cohort study using electronic records (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  word_document:
    toc: yes
    reference_docx: template.docx
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
link-citations: yes
always_allow_html: yes
---

@@

```{r, fig.width=10, fig.height=5}
knitr::include_graphics("study_diagram.svg")
```

@@

). We also compared care ratesDates of infectiondischarge were used alongside the death indicator to estimate the case-fatality rate in home and hospital, as well as compare all-cause mortality in care homes withbased on the presence of symptoms and without outbreaks, defined as at least one /or confirmed case or Covid-19 related death during the study period. infection. A time-dependent Cox proportional hazards model analysed the effect of the same factors on all-cause mortality based on the time-dependent infection classification (susceptible; symptomatic not confirmed; asymptomatic confirmed; symptomatic confirmed). We also compared rates of infection and mortality in care homes with and without outbreaks, defined as at least one confirmed infection or Covid-19 related death during the study period. A time-dependent Cox proportional hazards model analysed the effect of the same factors on all-cause mortality based on the time-dependent infection classification (uninfected; symptomatic not confirmed; asymptomatic confirmed; symptomatic confirmed) as well as the time-dependent home outbreak status (no Covid-related death or confirmed infection; at least one Covid-related death or confirmed infection). The associated hazard ratios enable us to compare the mortality in each group as compared to the reference category (uninfected in non-outbreak care homes). Interpretation is facilitated by computing the associated attributable risk $ AR = (HR - 1)/HR$, which

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5259851/pdf/12874_2016_Article_285.pdf

TO DOOOOO

@@

```{r fig_KM_CT_plots, fig.height = 4.5, fig.width = 14, message=FALSE}
write.csv(km_ct_residents_symp, file = "supp_table_kaplan_meier_residents_symptomatic.csv", row.names = F)
write.csv(km_ct_residents_confirmed, file = "supp_table_kaplan_meier_residents_confirmed.csv", row.names = F)
write.csv(km_ct_residents_deaths, file = "supp_table_kaplan_meier_residents_deaths.csv", row.names = F)

@@

km_residents_graph <- ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = Event), data = all_km_residents) +
  # facet_grid(region_group ~ .) +
  geom_ribbon(alpha=0.1, linetype = 2 ) +
  geom_path(lwd=1.2) +
  scale_y_continuous("Cumulative incidence",
                     breaks = seq(0,1,.02)) +
  ggsci::scale_color_lancet() +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(all_km_residents$date)),
               labels = format.Date(
                 unique(get_monday_date(all_km_residents$date)),
                 "%e %b %Y"
               )) +
  theme_bw()  +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1)) +
  coord_cartesian(ylim = c(0,0.27)) + ggtitle("(A) Residents")

@@

km_staff_graph <- ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = Event), data = all_km_staff) +
  # facet_grid(region_group ~ .) +
  geom_ribbon(alpha=0.1, linetype = 2,stat = ) +
  geom_path(lwd=1.2) +
  ggsci::scale_color_lancet()+
  scale_y_continuous("Cumulative incidence",
                     breaks = seq(0,1,.02)) +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(km_staff_symptomatic$date)),
               labels = format.Date(
                 unique(get_monday_date(km_staff_symptomatic$date)),
                 "%e %b %Y"
               )) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1)) +
  coord_cartesian(ylim = c(0,0.27)) +  ggtitle("(B) Staff")

km_ct_plots <- ggpubr::ggarrange(
  km_residents_graph,
  km_staff_graph
)

pdf("figure2.pdf", 14,4)
km_ct_plots
dev.off()
km_ct_plots
```

@@

```{r tab_CFR_age_sex, results="asis"}
residents_overview %>%
  mutate(Age = ageg, Sex = gender) %>% 
  group_by(Age, Sex) %>% 
  summarise(
    n = n_distinct(resident_id),
    resident_days = sum(resident_days),
    confirmed_infections = sum(confirmed == 1),
    deaths = sum(status == "Deceased"),
    deaths_confirmed = sum(status == "Deceased" & confirmed == 1)) %>% 
  mutate(
    cfr = tbrounding(deaths_confirmed/confirmed_infections*100),
    cfr_lower = tbrounding(epitools::binom.exact(deaths_confirmed, confirmed_infections)$lower*100),
    cfr_upper = tbrounding(epitools::binom.exact(deaths_confirmed, confirmed_infections)$upper*100),
    RR = relative_risk(deaths_confirmed, confirmed_infections - deaths_confirmed,
                       deaths - deaths_confirmed, n - confirmed_infections -(deaths - deaths_confirmed)),
    RR_SE = relative_risk_SE(deaths_confirmed, confirmed_infections - deaths_confirmed,
                       deaths - deaths_confirmed, n - confirmed_infections -(deaths - deaths_confirmed))
  ) %>%  ungroup() %>% 
  transmute(
    Age,
    Sex,
    N = formatbm(n),
    `Confirmed infections` = confirmed_infections,
    `Total deaths` = deaths,
    `Deaths in confirmed infections` = deaths_confirmed,
    `Case-fatality rate (%)` = paste0(cfr, " [", cfr_lower, "; ", cfr_upper, "]"),
    `Relative risk` = paste0(
      tbrounding(RR, 2), " [",
      tbrounding(exp(log(RR) - 1.96 * RR_SE), 2), "; ",
      tbrounding(exp(log(RR) + 1.96 * RR_SE), 2), "]"
    )
  ) %>% kable(align=c("l", "l", "r", "r", "r", "r", "r", "r"))
```

@@

```{r CPH_confirmed,   results="asis"}
# cph_conf <- coxph(Surv(time = start_int, time2 = stop_int +1, event = confirmed) ~ gender + ageg+home_type+user_type2+  staff_to_bed_ratio+bed_tot_category  +occ_bedroom_ratio_g6, data = residents_overview)
#  summary(cph_conf)
# cox.zph(cph_conf)
# survfit(Surv(time = start_int, time2 = stop_int +1, event = confirmed)~factor(gender), data = residents_overview) %>% ggsurvplot(ylim=c(.85,1), conf.int = T, legend = "right")
explanatory = c("gender", "ageg", "home_type", "user_type2",  
                "staff_to_bed_ratio", "bed_tot_category", "occ_bedroom_ratio_g6"  )
dependent = "Surv(time = start_int, time2 = stop_int +1, event = confirmed)"
CPH_confirmed <-  finalfit(residents_overview,
                           dependent, 
                           explanatory, column = T,
                           add_dependent_label = F)  
knitr::kable(CPH_confirmed,
             row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
```

@@

```{r residents_daily}

residents_daily <- tline %>% 
  filter(rday==1)  %>% 
  left_join(home_first_case) %>%
  mutate(outbreak = dplyr::case_when(
    date >= first_case_date ~ 1L,
    date < first_case_date ~ 0L,
    TRUE ~ 0L
  )) %>% 
  group_by(resident_id) %>% 
  mutate(death = case_when(
    status == "Deceased" & date == max(date) ~ 1L,
    TRUE ~ 0L
  )) %>% 
  ungroup() %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1)) %>% 
  mutate(covid_status_outbreak = as.factor(case_when(
    symptomatic==0 & confirmed == 0 & outbreak == 0 ~ "0 Uninfected (other home)",
    symptomatic==1 & confirmed == 0 & outbreak == 0 ~ "1 Symptomatic not confirmed  (other home)",
    symptomatic==0 & confirmed == 0 & outbreak == 1 ~ "2 Uninfected (outbreak home)",
    symptomatic==1 & confirmed == 0 & outbreak == 1 ~ "3 Symptomatic not confirmed  (outbreak home)",
    symptomatic==0 & confirmed == 1 ~ "4 Confirmed asymptomatic",
    symptomatic==1 & confirmed == 1 ~ "5 Confirmed symptomatic"
  )))
residents_daily$covid_status_outbreak_digit <- as.factor(
  as.numeric(residents_daily$covid_status_outbreak)-1)
residents_daily <- bind_cols(
  residents_daily,
  as_tibble(model.matrix(~covid_status_outbreak_digit, residents_daily)[,-1])
)

# residents_daily <- residents_daily %>% 
#   group_by(date) %>% 
#   summarise(n = sum(rday),
#             d = sum(death),
#             covid_status_outbreak_digit0 = 1 - 
#               mean(covid_status_outbreak_digit1) -
#               mean(covid_status_outbreak_digit2) -
#               mean(covid_status_outbreak_digit3) -
#               mean(covid_status_outbreak_digit4) -
#               mean(covid_status_outbreak_digit5),
#             covid_status_outbreak_digit1 = mean(covid_status_outbreak_digit1),
#             covid_status_outbreak_digit2 = mean(covid_status_outbreak_digit2),
#             covid_status_outbreak_digit3 = mean(covid_status_outbreak_digit3),
#             covid_status_outbreak_digit4 = mean(covid_status_outbreak_digit4),
#             covid_status_outbreak_digit5 = mean(covid_status_outbreak_digit5)) %>% 
#   mutate(date_int = as.integer(date - date_min))

@@

```{r risk_death, results="asis"}
# cph_death2 <- coxph(Surv(date1, date2, death_status) ~
#                       gender+ ageg + home_type + covid_status_outbreak +
#                       staff_to_bed_ratio + bed_tot_category + occ_bedroom_ratio_g6 +
#                       cluster(resident_id), data = residents_episodes_short, ties = "breslow")
# summary(cph_death2)
# cox.zph(cph_death2)

@@

```{r, warning=F, results = "asis"}
cph_death2 <- coxph(Surv(date1, date2, death_status) ~
                      gender+ ageg + home_type + covid_status_outbreak +
                      staff_to_bed_ratio + bed_tot_category + occ_bedroom_ratio_g6 +
                      cluster(resident_id), data = residents_episodes_short)
HRs_covid <- tibble::enframe(cph_death2$coefficients[grep("covid_status_outbreak", 
                                                          names(cph_death2$coefficients))]) %>% 
  transmute(covid_status_outbreak = gsub("covid_status_outbreak", "", name),
            HR = exp(value))
HRs_CI_covid <- as_tibble(exp(confint(cph_death2))[grep("covid_status_outbreak", 
                                                          names(cph_death2$coefficients)),]) %>% 
  rename(HR_lower =  `2·5 %`,
         HR_upper = `97·5 %`)
HRs_CI_covid$covid_status_outbreak <- gsub(
  "covid_status_outbreak", "",
  grep("covid_status_outbreak", names(cph_death2$coefficients), value = T)
) 
HRs_covid <- merge(HRs_covid, HRs_CI_covid)

attributable_deaths <- residents_episodes_short %>% 
  filter(death_status == 1) %>% 
  group_by(covid_status_outbreak) %>% 
  summarise(Deaths = n()) %>% 
    left_join(HRs_covid) %>% 
    mutate(HR = tidyr::replace_na(HR, 1),
           HR_lower = tidyr::replace_na(HR_lower, 1),
           HR_upper = tidyr::replace_na(HR_upper, 1)) %>% 
    mutate(AH = (HR - 1)/HR,
           AH_lower = (HR_lower - 1)/HR_lower,
           AH_upper = (HR_upper - 1)/HR_upper) %>% 
    mutate(`Attributable deaths` = AH * Deaths) %>% 
    mutate(`AD (%)` = tbrounding(`Attributable deaths`/ sum(Deaths)*100)) %>% 
  transmute(`Infection/outbreak status` = covid_status_outbreak,
            Deaths,
            HR = paste0(tbrounding(HR,2), " [",
                        tbrounding(HR_lower,2),"; ",
                        tbrounding(HR_upper,2), "]"),
            `Attibutable hazard (%)`= paste0(tbrounding(AH*100,1), " [",
                        tbrounding(AH_lower*100,1),"; ",
                        tbrounding(AH_upper*100,1), "]"),
           `Attributable deaths` = round(`Attributable deaths`),
           `AD (%)`)
attributable_deaths <- bind_rows(
  attributable_deaths,
  tibble(
    `Infection/outbreak status` = "TOTAL",
            Deaths = (sum(attributable_deaths$Deaths)),
            HR = "",
            `Attibutable hazard (%)`= "",
           `Attributable deaths` = sum(attributable_deaths$`Attributable deaths`),
           `AD (%)` = tbrounding(
             sum(attributable_deaths$`Attributable deaths`)/sum(attributable_deaths$Deaths)*100)
  )
)

kable(attributable_deaths, align =c("l", "r", "r", "r", "r", "r", "r"))
```

```{r}

# residents_episodes$resident_id_int <- as.numeric(as.factor(residents_episodes$resident_id))
# cph_death2 <- coxph(Surv(date1, date2, death_status) ~
#                       gender+ ageg + home_type + covid_status_outbreak_digit1 +
#                       covid_status_outbreak_digit2 + covid_status_outbreak_digit3+
#                       covid_status_outbreak_digit4 +covid_status_outbreak_digit5 +
#                       staff_to_bed_ratio + bed_tot_category + occ_bedroom_ratio_g6 +
#                       cluster(resident_id_int), data = data.frame(residents_episodes), 
#                     ties = "breslow")
# 
# trucmuche <- survfit(Surv(date1, date2, death_status) ~
#                       gender+ ageg + home_type + covid_status_outbreak_digit1 +
#                       covid_status_outbreak_digit2 + covid_status_outbreak_digit3+
#                       covid_status_outbreak_digit4 +covid_status_outbreak_digit5 +
#                       staff_to_bed_ratio + bed_tot_category + occ_bedroom_ratio_g6,
#                      residents_episodes_short, id = "resident_id")
# 
# AH_fractions <- residents_daily %>% 
#   tidyr::pivot_longer(cols=starts_with("covid_status_outbreak_")) %>% 
#   dplyr::rename(prevalence = value) %>% 
#   left_join(tibble::enframe(cph_death2$coefficients[grep("covid_status_outbreak_digit", names(cph_death2$coefficients))])) %>% 
#   mutate(AH = (prevalence * tidyr::replace_na(exp(value) - 1, 0)) / 
#            (prevalence * tidyr::replace_na(exp(value) - 1, 0) +1))
# 
# 
# AH_coefs <- tibble::enframe(cph_death2$coefficients[grep("covid_status_outbreak_digit",
#                                                          names(cph_death2$coefficients))]) %>% 
#   rename(covid_status_outbreak_digit = name) %>% 
#   mutate(covid_status_outbreak_digit = gsub("covid_status_outbreak_digit", "", covid_status_outbreak_digit))
#                               
# AH_fractions <- residents_daily %>%
#   group_by(date, covid_status_outbreak_digit) %>%
#   summarise(n = sum(rday),
#             d = sum(death)) %>% 
#   left_join(AH_coefs)  %>% 
#   filter(date >= as.Date("2020-04-01")) %>% 
#   mutate(AH =  (tidyr::replace_na(exp(value), 1) -1 ) / tidyr::replace_na(exp(value), 1)) %>% 
#   group_by(covid_status_outbreak_digit) %>% 
#   summarise(AH = sum(AH*d),
#             d = sum(d),
#             n = sum(n))  
#             
# 
# 
#  
# 
# ggplot(AH_fractions)+
#   geom_line(aes(x = date, y = AH*d, color = name ))
#

# 
# # summary(cph_death2)
# AF_estimates <- list()
# for(i in 1:5) {
#   AF_estimates[[levels(residents_episodes$covid_status_outbreak)[i]]] <- 
#     AF::AFcoxph(object = cph_death2, data = data.frame(residents_episodes), 
#             exposure = paste0("covid_status_outbreak_digit", i), clusterid = "resident_id_int")
#   AF_estimates[[i]]$se <- sqrt(AF_estimates[[i]]$AF.var)
#   AF_estimates[[i]]$confidence.interval <- AF:::CI.AF(
#     AF = AF_estimates[[i]]$AF.est, 
#     Std.Error = AF_estimates[[i]]$se, 
#     confidence.level = 0.95, CI.transform = "log")
#   colnames(AF_estimates[[i]]$confidence.interval) <- c("Lower limit", "Upper limit")
#   AF_estimates[[i]]$covid_status_outbreak <- levels(
#     residents_episodes$covid_status_outbreak)[i+1]
#   AF_estimates[[i]] <- bind_cols(
#     as_tibble(
#       AF_estimates[[i]][c('times', 'covid_status_outbreak', 'AF.est', 'AF.var', 
#                          'S.est', 'S0.est', 'S.var', 'S0.var')]
#     ),
#     as_tibble(AF_estimates[[i]]$confidence.interval)
#   )
# }
# 
# 
# ggplot(bind_rows(AF_estimates[2:5]),
#        aes(x = times,  color =  covid_status_outbreak )) +
#   geom_ribbon(aes(ymin = `Lower limit`*100, ymax = `Upper limit`*100), 
#               alpha=0.05, linetype = 2) +
#   geom_line(aes(y=AF.est*100), size = 1) +
#   coord_cartesian(ylim=c(0, 30)) +
#   ggsci::scale_color_lancet() +
#   xlab("Excess mortality (%)") +
#   ggtitle("% excess mortality compared with uninfected in non-outbreak homes")
# 
# # 
# cph_death2 <- coxph(Surv(date1, date2, death_status) ~
#                         covid_status_outbreak_digit  +
#                       cluster(resident_id_int), data = residents_episodes, 
#                     ties = "breslow")
# cph_death2 <- flexsurv::flexsurvreg(Surv(date1, date2, death_status) ~
#                         covid_status_outbreak_digit  +
#                       cluster(resident_id), 
#                       data = residents_episodes_short,dist = "weibull")
# 
# truc <- survfit(cph_death2, tibble(date1 = 30, covid_status_outbreak_digit = factor(1:5, levels = 1:5) )) 
#   ggsurvplot(truc,
#              data = tibble(date1 = 30, covid_status_outbreak_digit = factor(5, levels = 1:5) ))
#     
```

>>>>

---
title: "Covid-19 infection and attributable mortality in residents of UK care homes: Cohort study using electronic records (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
  word_document:
    toc: yes
    reference_docx: template.docx
link-citations: yes
always_allow_html: yes
---

@@

```{r study_diagram, fig.width=10, fig.height=5}
knitr::include_graphics("study_diagram.svg")
```

@@

). We also compared care ratesDates of infectiondischarge were used alongside the death indicator to estimate the case-fatality rate in home and hospital, as well as compare all-cause mortality in care homes withbased on the presence of symptoms and without outbreaks, defined as at least one /or confirmed case or Covid-19 related death during the study period. infection. A time-dependent Cox proportional hazards model analysed the effect of the same factors on all-cause mortality based on the time-dependent infection classification (susceptible; symptomatic not confirmed; asymptomatic confirmed; symptomatic confirmed). We also compared rates of infection and mortality in care homes with and without outbreaks, defined as at least one confirmed infection or Covid-19 related death during the study period. A time-dependent Cox proportional hazards model analysed the effect of the same factors on all-cause mortality based on the time-dependent infection classification (uninfected; symptomatic not confirmed; asymptomatic confirmed; symptomatic confirmed) as well as the time-dependent home outbreak status (no Covid-related death or confirmed infection; at least one Covid-related death or confirmed infection). The associated hazard ratios enable us to compare the mortality in each group as compared to the reference category (uninfected in non-outbreak care homes). The corresponding attributable fraction is obtained by using the model to predict the counterfactual mortality, then computing AFW, the attributable fraction within study:[@Samuelsen2008]

$$ AFW = \dfrac{\sum_{i=1}^{n}(\hat{F}_i - \hat{F}_i^\star)}{\sum_{i=1}^{n}\hat{F}_i} $$

where $\hat{F}_i$ designates the predicted probability of death of individual $i = \{1, \dots, n\}$ based on the observed infection/outbreak classification, and  $\hat{F}_i^\star$ the same probability had the individual been uninfected and living in a non-outbreak care home. Observed censoring times are used.

@@

```{r fig_KM_CT_plots, fig.height = 4, fig.width = 12, message=FALSE}
write.csv(km_ct_residents_symp, file = "supp_table_kaplan_meier_residents_symptomatic.csv", row.names = F)
write.csv(km_ct_residents_confirmed, file = "supp_table_kaplan_meier_residents_confirmed.csv", row.names = F)
write.csv(km_ct_residents_deaths, file = "supp_table_kaplan_meier_residents_deaths.csv", row.names = F)

@@

km_residents_graph <- ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = Event), data = all_km_residents) +
  geom_ribbon(alpha=0.1, linetype = 2 ) +
  geom_path(lwd=1.2) +
  geom_text(aes(label = if_else(
    date == get_monday_date(date),
    formatbm(round((1-S_t) * desc_residents$CT_total_residents)),
    ""
  )), position = position_nudge(y = .02)) +
  scale_y_continuous("Cumulative incidence",
                     breaks = seq(0,1,.02)) +
  ggsci::scale_color_lancet() +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(all_km_residents$date)),
               labels = format.Date(
                 unique(get_monday_date(all_km_residents$date)),
                 "%e %b %Y"
               )) +
  theme_bw()  +
  theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1)) +
  coord_cartesian(ylim = c(0,0.27)) + ggtitle("(A) Residents")

@@

km_staff_graph <- ggplot(aes(x = date, y = 1-S_t, 
           ymax = 1 - St_max,
           ymin = 1 - St_min,
           color = Event), data = all_km_staff) +
  geom_ribbon(alpha=0.1, linetype = 2,stat = ) +
  geom_path(lwd=1.2) +
  geom_text(aes(label = if_else(
    date == get_monday_date(date),
    formatbm(round((1-S_t) * desc_staff$total_staff)),
    ""
  )), position = position_nudge(y = .02)) +
  ggsci::scale_color_lancet()+
  scale_y_continuous("Cumulative incidence",
                     breaks = seq(0,1,.02)) +
  scale_x_date(name = "Date", 
               breaks = unique(get_monday_date(km_staff_symptomatic$date)),
               labels = format.Date(
                 unique(get_monday_date(km_staff_symptomatic$date)),
                 "%e %b %Y"
               )) +
  theme_bw() +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 25, vjust = 1, hjust=1)) +
  coord_cartesian(ylim = c(0,0.27)) +  ggtitle("(B) Staff")

km_ct_plots <- ggpubr::ggarrange(
  km_residents_graph,
  km_staff_graph,
  widths = c(1.4, 1)
)

pdf("figure2.pdf", 12,4)
km_ct_plots
dev.off()
km_ct_plots
```

@@

```{r tab_CFR_age_sex, results="asis"}
residents_overview %>%
  mutate(Age = ageg, Sex = gender) %>% 
  group_by(Sex, Age) %>% 
  summarise(
    n = n_distinct(resident_id),
    resident_days = sum(resident_days),
    confirmed_infections = sum(confirmed == 1),
    deaths = sum(status == "Deceased"),
    deaths_confirmed = sum(status == "Deceased" & confirmed == 1)) %>% 
  mutate(
    cfr = tbrounding(deaths_confirmed/confirmed_infections*100),
    cfr_lower = tbrounding(epitools::binom.exact(deaths_confirmed, confirmed_infections)$lower*100),
    cfr_upper = tbrounding(epitools::binom.exact(deaths_confirmed, confirmed_infections)$upper*100),
    RR = relative_risk(deaths_confirmed, confirmed_infections - deaths_confirmed,
                       deaths - deaths_confirmed, n - confirmed_infections -(deaths - deaths_confirmed)),
    RR_SE = relative_risk_SE(deaths_confirmed, confirmed_infections - deaths_confirmed,
                       deaths - deaths_confirmed, n - confirmed_infections -(deaths - deaths_confirmed))
  ) %>%  ungroup() %>% 
  transmute(
    Age,
    Sex,
    N = formatbm(n),
    `Confirmed infections` = confirmed_infections,
    `Total deaths` = deaths,
    `Deaths in confirmed infections` = deaths_confirmed,
    `Case-fatality rate (%)` = paste0(cfr, " [", cfr_lower, "; ", cfr_upper, "]")
    # `Relative risk` = paste0(
    #   tbrounding(RR, 2), " [",
    #   tbrounding(exp(log(RR) - 1.96 * RR_SE), 2), "; ",
    #   tbrounding(exp(log(RR) + 1.96 * RR_SE), 2), "]"
    # )
  ) %>% kable(align=c("l", "l", "r", "r", "r", "r", "r", "r"))
```

@@

```{r hazard_ratios_confirmed,   results="asis"}
# cph_conf <- coxph(Surv(time = start_int, time2 = stop_int +1, event = confirmed) ~ gender + ageg+home_type+user_type2+  staff_to_bed_ratio+bed_tot_category  +occ_bedroom_ratio_g6, data = residents_overview)
#  summary(cph_conf)
# cox.zph(cph_conf)
# survfit(Surv(time = start_int, time2 = stop_int +1, event = confirmed)~factor(gender), data = residents_overview) %>% ggsurvplot(ylim=c(.85,1), conf.int = T, legend = "right")

explanatory = c("gender", "ageg", "home_type", "user_type2",  
                "staff_to_bed_ratio", "bed_tot_category", "occ_bedroom_ratio_g6"  )
dependent = "Surv(time = start_int, time2 = stop_int +1, event = confirmed)"
CPH_confirmed <-  finalfit(residents_overview,
                           dependent, 
                           explanatory, column = T,
                           add_dependent_label = F)  
knitr::kable(CPH_confirmed,
             row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
```

@@

```{r residents_daily}
residents_daily <- tline %>% 
  filter(rday==1)  %>% 
  left_join(home_first_case) %>%
  mutate(outbreak = dplyr::case_when(
    date >= first_case_date ~ 1L,
    date < first_case_date ~ 0L,
    TRUE ~ 0L
  )) %>% 
  group_by(resident_id) %>% 
  mutate(death = case_when(
    status == "Deceased" & date == max(date) ~ 1L,
    TRUE ~ 0L
  )) %>% 
  ungroup() %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1)) %>% 
  mutate(covid_status_outbreak = as.factor(case_when(
    symptomatic==0 & confirmed == 0 & outbreak == 0 ~ "0 Uninfected (other home)",
    symptomatic==1 & confirmed == 0 & outbreak == 0 ~ "1 Symptomatic not confirmed  (other home)",
    symptomatic==0 & confirmed == 0 & outbreak == 1 ~ "2 Uninfected (outbreak home)",
    symptomatic==1 & confirmed == 0 & outbreak == 1 ~ "3 Symptomatic not confirmed  (outbreak home)",
    symptomatic==0 & confirmed == 1 ~ "4 Confirmed asymptomatic",
    symptomatic==1 & confirmed == 1 ~ "5 Confirmed symptomatic"
  )))
residents_daily$covid_status_outbreak_digit <- as.factor(
  as.numeric(residents_daily$covid_status_outbreak)-1)
residents_daily <- bind_cols(
  residents_daily,
  as_tibble(model.matrix(~covid_status_outbreak_digit, residents_daily)[,-1])
)

@@

```{r hazard_ratios_death, results="asis"}
# cph_death2 <- coxph(Surv(date1, date2, death_status) ~
#                       gender+ ageg + home_type + covid_status_outbreak +
#                       staff_to_bed_ratio + bed_tot_category + occ_bedroom_ratio_g6 +
#                       cluster(resident_id), data = residents_episodes_short, ties = "breslow")
# summary(cph_death2)
# cox.zph(cph_death2)

@@

```{r attributable_deaths, warning=F, message=F, results = "asis"}
cph_death3 <- coxph(Surv(date1, date2, death_status) ~
                    gender+ ageg + home_type + covid_status_outbreak_digit +
                    staff_to_bed_ratio + bed_tot_category + occ_bedroom_ratio_g6 +
                    cluster(resident_id),
                    data =  residents_episodes_short)

HRs_covid <- tibble::enframe(cph_death3$coefficients[grep("covid_status_outbreak", 
                                                          names(cph_death3$coefficients))]) %>% 
  transmute(covid_status_outbreak_digit = gsub("covid_status_outbreak_digit", "", name),
            HR = exp(value))
HRs_CI_covid <- as_tibble(exp(confint(cph_death3))[grep("covid_status_outbreak", 
                                                          names(cph_death3$coefficients)),]) %>% 
  rename(HR_lower =  `2·5 %`,
         HR_upper = `97·5 %`)
HRs_CI_covid$covid_status_outbreak_digit <- gsub(
  "covid_status_outbreak_digit", "",
  grep("covid_status_outbreak", names(cph_death3$coefficients), value = T)
) 
HRs_covid <- merge(HRs_covid, HRs_CI_covid)

attributable_deaths <- residents_episodes_short %>%
  group_by(covid_status_outbreak_digit, covid_status_outbreak) %>%
  summarise(
    rdays = sum(as.integer(date2-date1)),
    deaths = sum(death_status == 1)
    ) %>% 
  left_join(HRs_covid) %>% 
  mutate(HR = tidyr::replace_na(HR, 1),
           HR_lower = tidyr::replace_na(HR_lower, 1),
           HR_upper = tidyr::replace_na(HR_upper, 1))

death_predictions <- 1-predict(cph_death3, residents_episodes_short, type = "survival")

att_fractions <- 0
for(i in 1:5) {
  residents_episodes_short_counterfact <- mutate(
    residents_episodes_short, 
    # date2 = if_else(death_status == 1 & covid_status_outbreak_digit == as.character(i), 105, date2),
    # death_status = if_else(death_status == 1 & 
    #                          covid_status_outbreak_digit == as.character(i),
    #                        0L, death_status),
    covid_status_outbreak_digit = case_when(
      covid_status_outbreak_digit == as.character(i) ~ factor("0", levels = 0:5),
      TRUE ~ covid_status_outbreak_digit)
  )
  death_predictions_counterfactual <- 1 - predict(cph_death3, 
                                                  residents_episodes_short_counterfact, 
                                                  type = "survival")
  att_fractions[i+1] <- sum(death_predictions - death_predictions_counterfactual) / sum(death_predictions)
}

rm(death_predictions, death_predictions_counterfactual)

attributable_deaths$AF <- att_fractions
attributable_deaths <- bind_rows(attributable_deaths, 
                                 colSums(attributable_deaths[,c("rdays", "deaths", "AF")]))
attributable_deaths %>% ungroup() %>% transmute(
  `Infection/outbreak status` = tidyr::replace_na(as.character(covid_status_outbreak), "TOTAL"),
  `Ajusted HR` = paste0(tbrounding(HR,2), " [",
                        tbrounding(HR_lower,2),"; ",
                        tbrounding(HR_upper,2), "]"),
  `Resident-days` = formatbm(rdays),
  `Total deaths` = formatbm(deaths),
  `Deaths attributable to Covid-19 (% all-cause deaths)` = paste0(
    round(AF * 1330), " (",
    tbrounding(AF*100, 2), ")")
) %>% kable(align=c("l", rep("r", 4)))
